---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Loading and preprocessing the data (also unzip raw data)
```{r}
# set working directory to "/Users/brad/Documents/development/repos/coursera/reproducible research/project 1"
setwd("/Users/brad/Documents/development/repos/_Coursera/reproducible research/project 1/RepData_PeerAssessment1")

# Download the raw data file (zipped)
# Raw File location: https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip

raw_file <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
curr_dir <- getwd() 
raw_file_full_path <- paste(getwd(),"/raw data/repdata_data_activity.zip",sep="")

download.file(url = raw_file,
              destfile= raw_file_full_path, 
              method="curl")

# Extract the zip file
# unzip the data file and rename it locally
unzip(zipfile = raw_file_full_path, exdir=".//raw data")

# read in the data
activity <- read.csv("./raw data/activity.csv",header = TRUE)
summary(activity)
```

## Question 1: What is mean total number of steps taken per day?
### Calculate total number of steps per day
```{r}
steps_per_day <- tapply(X = activity$steps, INDEX = activity$date, FUN = sum)
```
### Make a histogram of the total number of steps taken each day
```{r}
hist(steps_per_day,
     col = 2,
     main ="Total Steps Per Day (Histogram)",
     ylim = c(0,30),
     xlab = "Number of Steps", 
     ylab = "Frequency")
```

### Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean_steps_per_day <- mean(steps_per_day, na.rm = TRUE)
median_steps_per_day <- median(steps_per_day, na.rm = TRUE)
paste("Mean number of steps per day: ",mean_steps_per_day)
paste("Median number of steps per day: ",median_steps_per_day)
```
## Question 2: What is the average daily activity pattern?

### Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
total_steps_per_interval <- tapply(activity$steps, activity$interval, mean, na.rm = TRUE)
plot(as.numeric(names(total_steps_per_interval)), 
     total_steps_per_interval, 
     main = "Average Daily Activity Pattern",
     xlab = "Interval",
     ylab = "# Steps", 
     col = 4,
     type = "l")
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
# Sort the total steps per interval in decreasing order and grab the top result.
max_interval <- names(sort(total_steps_per_interval, decreasing = TRUE)[1])
max_steps <- sort(total_steps_per_interval, decreasing = TRUE)[1]
paste("The interval with the maximum number of steps is",max_interval)
paste("The maximum number of steps in that interval is",max_steps)
```

## Question 3: Imputing missing values
### Calculate and report the total number of missing values in the dataset.
```{r}
num_na_values <- sum(is.na(activity$steps))
paste("The number of NA values is",num_na_values)
```
### Devise a strategy for filling in all of the missing values in the dataset. 
```{r}
```
### Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
```
### Make a histogram of the total number of steps taken each day and calculate and report the mean and median total number of steps taken per day. 
```{r}
```
## Question 4: Are there differences in activity patterns between weekdays and weekends?
### Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
```
### Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
```
